Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-04 inc/hardware/laser/laser.tex
more tweaks to laser section
merge additional comments on laser section

Since the initial construction and characterization of this laser system, described above, several improvements have been made, primarily to improve the long-term stability of the laser.
Due to the laser's role as a component of the larger UEM system, its stability is a tremendously important factor, both for repeatability of measurements and suitability for use in a user instrument.

The second harmonic generation system now employs a 3mm non-critically phase matched LBO crystal.
This change has two benefits. 
First, in this geometry, the walk-off angle between the fundamental and second harmonic is zero, while the group velocity mismatch of 150fs ($\sim$ 50 fs/mm) is much less than the pulse duration 250fs. %TODO cite Koechner?
Second, because in this case, non-critically phase matching requires that the crystal is held at an elevated temperature, well above the boiling point of water, the deleterious absorption of water in the hygroscopic LBO crystal is prevented.
Note that additionally, in the place of the curved mirrors originally employed, the second harmonic system now uses 75mm lenses to focus the laser into the crystal and collimate the output.

Although the laser system had been relatively robust, maintaining mode-locking operation was initially problematic.
A cover was added above the laser cavity to prevent dust and air currents from interrupting mode locking.
More importantly, a 1.4mm aperture (approximately four times the HW1/eM beam width at the output coupler) was added to suppress higher-order transverse modes.
As this is an additional loss in the cavity, the output coupler was changed from 93\% to 95\% reflectivity to maintain the soliton modelocking power.

These changes have greatly improved the mode locking stability.
The laser now regularly operates at $1.9 \pm 0.1$W in mode locked operation with 250fs pulse durations, with minimal regular maintenance.
The 40-50\% efficient second harmonic generation yields 0.8-1W of green output.
The green radiation can be again doubled to the ultraviolet (UV) in a 6mm $\beta$-Barium borate (BBO) crystal, though the group velocity mismatch results in a 4ps pulse duration.
All data presented in this thesis has been acquired subsequent to these performance enhancements.

>>>>

Since the initial construction and characterization of this laser system, described above, several improvements have been made, primarily to improve the long-term mode-locking stability of the laser.
Due to the laser's role as a component of the larger UEM system, its stability is a tremendously important factor, both for repeatability of measurements and suitability for use in a user instrument.

The second harmonic generation system now employs a 3mm non-critically phase matched LBO crystal.
This change has two benefits. 
First, in this geometry, the walk-off angle between the fundamental and second harmonic is zero, while the group velocity mismatch of 150fs ($\sim$ 50 fs/mm) is still less than the pulse duration 250fs. %TODO cite Newlight Photonics, Canada
Second, because in this case, non-critically phase matching requires that the crystal is held at an elevated temperature, well above the boiling point of water, the deleterious absorption of water in the hygroscopic LBO crystal is prevented.
%TODO replace above with once oven temperature is known
%Second, because in this case, non-critically phase matching requires that the crystal is held at an elevated temperature of XXX$^{\circ}$C, the deleterious absorption of water in the hygroscopic LBO crystal is prevented.
Additionally, in the place of the curved mirrors originally employed, the second harmonic system now uses 75mm lenses to focus the laser into the crystal and collimate the output.

Although the laser system had been relatively robust, maintaining mode-locking operation was initially problematic.
A cover was added above the laser cavity to prevent dust and air currents from interrupting mode locking.
More importantly, a 1.4mm aperture (approximately four times the HW1/eM beam width at the output coupler) was added to suppress higher-order transverse modes.
As this is an additional loss in the cavity, the output coupler was changed from 93\% to 95\% reflectivity to maintain the soliton modelocking power.
These changes have greatly improved the mode locking stability.

The laser now regularly operates at $1.9 (\pm 0.1)$W in mode locked operation with 250fs pulse durations, with minimal regular maintenance.
The 40-50\% efficient second harmonic generation yields 0.8-1W of green output.
The green radiation can be again doubled to yield 261nm ($\hbar \omega = 4.75$eV) ultraviolet (UV) radiation in a 6mm $\beta$-Barium borate (BBO) crystal, though the group velocity mismatch results in a 4ps pulse duration.
All data presented in this thesis has been acquired subsequent to these performance enhancements.
